---
format: live-html
engine: knitr
lightbox: true
---
#  Boxplot from file

From the previous pages, we have a dataset with three columns:

| Column | Data |
|------------------------------------|------------------------------------|
| serotoninLevel | Measurement of serotonin in individual locusts |
| treatmentTime | Amount of time being crowded (0, 1 or 2 hours) |
| serotoninCategory | Rank of serotonin level in three categories (High, Medium, Low) |

: Locust Serotonin data columns and descriptions {.striped .hover}

Next, we're going to use boxplots to compare the serotonin level for different treatment groups.

```{r}
#| eval: false
# Load library
library(ggplot2)
# Import the CSV file into an R Dataframe
locust_df <- read.csv("LocustSerotonin.csv") 

# Plot the data using ggplot package
ggplot(locust_df, aes(x = treatmentTime, 
                      y = serotoninLevel, 
                      group = treatmentTime)) +
  geom_boxplot() 
```
```{r}
#| echo: false
# Load library
library(ggplot2)
# Import the CSV file into an R Dataframe
locust_df <- read.csv("../Datasets/LocustSerotonin.csv") 

# Plot the data using ggplot package
ggplot(locust_df, aes(x = treatmentTime, 
                      y = serotoninLevel, 
                      group = treatmentTime)) +
  geom_boxplot() 
```

::: {.callout-note title="Question"}
What happens if you remove `group = treatmentTime` from the above code? What is this doing?
:::

Let's break that down:

-   `ggplot(locust_df, aes(x = treatmentTime, y = serotoninLevel, group = treatmentTime))`:
    -   `ggplot()` is the graphing function we are using
    -   `locust_df` is the data we want to plot
    -   `aes()` is our *aesthetics* - what is our x and y values?
    -   **group**: Our treatmentTime column is a continuous data (as it is numerical). We want this to be discrete (three separate groups) so we explicity tell ggplot to organise this data into groups.
-   `geom_boxplot()`: this is specifying we want to plot a box plot of our data
