---
format: live-html
engine: knitr
lightbox: true
---

{{< include /_extensions/r-wasm/live/_knitr.qmd >}}

# R functions

We won't be writing our own R functions, but we'll be using in-built functions and functions from libraries.

A function is like a mini-program - it's a set of code that somebody else has written that we can apply to our own variables.

These work using the following structure: `function_name(arguments)` where:

-   `function_name` is the name of the function
-   `arguments` are the values we are putting in, or any options we want to change

See examples below using some [in-built functions](https://www.javatpoint.com/r-built-in-functions):

```{webr}
# Calculate mean and standard deviation (sd)
my_numbers <- c(3, 7, 9, 12)
mean(my_numbers) # calculate the mean 
sd(my_numbers) # calculate the standard deviation

# Convert string to lower case 
my_string <- "HELLO WORLD"
tolower(my_string)

```

We can also give *multiple arguments* by separating them with commas.

In the example below, we're using the `seq()` function to generate a sequence of numbers. We want the sequence to start at 0, end at 20, and go up by 2 every number.

```{webr}
# Generate a range of numbers 
mysequence <- seq(0, 20, by=2)
mysequence
```

### Give it a go:

From what you've learnt so far, can you:

1.  Generate every third number from 1 to 100 (1, 3, 6...)
2.  Find the mean of those numbers
3.  Find the standard deviation

```{webr}
#| exercise: ex_1
# Write your answer here
#
#
```

:::: {.hint exercise="ex_1"}
::: {.callout-note collapse="false"}
## Hint 1

Consider using the `seq()` function

```{r}
#| eval: false
seq(start value, stop value, by=increase)
```
:::
::::

:::: {.hint exercise="ex_1"}
::: {.callout-note collapse="false"}
## Hint 2

Consider using the functions `seq()`, `mean()` and `sd()`

```{r}
#| eval: false
variablename <- seq(start value, stop value, by=increase)
mean(variablename)
sd(variablename)
```
:::
::::

::: {.callout-tip title="Show answer" collapse="true"}
```{webr}
# Generate a range of numbers 
mysequence <- seq(1, 100, by=3)
mysequence
mean(mysequence) # calculate the mean 
sd(mysequence) # calculate the standard deviation
```
:::
