---
format: live-html
engine: knitr
lightbox: true
webr:
  resources:
    - https://raw.githubusercontent.com/LizzieWadsworth/Level-4-R-course-book/refs/heads/main/Datasets/LocustSerotonin.csv
  packages:
    - ggplot2
    - dplyr
---

{{< include /_extensions/r-wasm/live/_knitr.qmd >}}

# Unpaired t-test

We're going to be using the same examples as in Statistics lecture 3 (by Dr Trigiante) for each of our tests. For this example, we'll use the Diabetic and Non-Diabetic glycemia values. 

## Add data to R 

The first thing we need to do is copy our data in R. As these are fairly small datasets, we can manually type in our data using the `c()` function to create a vector of values.

`c()` is short for combine, and is used for combining sets of values (in this case, numbers) into an array (data structure)

```{r}
# Create a vector of values for each of our columns
NonDiabetic = c(82,87,79,79,86,70,98,92,75,90,73,70,87,88,96,93) # NonDiabetic glycemia values
Diabetic = c(112,110,94,121,85,125,80,122,117,116,99,90,99,95,84,84) # Diabetic glycemia values
```

## Calculate mean and interquartile range

We can use the `summary()` function to view our mean and interquartile range.

```{r}
# Use summary() function to see the Interquartile Range
summary(NonDiabetic) # print summary of NonDiabetic glycemia values to the console
summary(Diabetic) # print summary of Diabetic glycemia values to the console
```

## Calculate unpaired T test

Using the function `t.test()`, we can add our two sets of values as arguments and calculate the p-value. This should match the result from excel of an **unpaired** t-test. 

```{r}
# Run an unpaired t test using t.test()
unpaired_t_test_result <- t.test(NonDiabetic, Diabetic)
# Print the result to the console
unpaired_t_test_result
```

That's our t-test done already!