---
format: live-html
engine: knitr
lightbox: true
webr:
  resources:
    - https://raw.githubusercontent.com/LizzieWadsworth/Level-4-R-course-book/refs/heads/main/Datasets/LocustSerotonin.csv
  packages:
    - ggplot2
    - dplyr
---

{{< include /_extensions/r-wasm/live/_knitr.qmd >}}

# Unpaired t-test

We're going to be using the same examples as in Statistics lecture 3 (by Dr Trigiante) for each of our tests. For this example, we'll use the Diabetic and Non-Diabetic glycemia values. 

The first thing we need to do is copy our data in R. As these are fairly small datasets, we can manually type in our data using the `c()` function to create a vector of values.

```{r}
# Create a vector of values for each of our columns
NonDiabetic = c(82,87,79,79,86,70,98,92,75,90,73,70,87,88,96,93) # NonDiabetic glycemia values
Diabetic = c(112,110,94,121,85,125,80,122,117,116,99,90,99,95,84,84) # Diabetic glycemia values
```

We can use the `summary()` function to view our mean and interquartile range.

```{r}
# Use summary() function to see the Interquartile Range
summary(NonDiabetic) # print summary of NonDiabetic glycemia values to the console
summary(Diabetic) # print summary of Diabetic glycemia values to the console
```

Using the function `t.test()`, we can add our two sets of values as arguments and calculate the p-value. This should match the result from excel of an **unpaired** t-test. 

```{r}
# Run an unpaired t test using t.test()
unpaired_t_test_result <- t.test(NonDiabetic, Diabetic)
# Print the result to the console
unpaired_t_test_result
```

That's our t-test done already!

## T-test options

We didn't have to use any **options** here as the default test is two tailed, unpaired, and assuming the variance is not equal. If you want to change any of the options, you can try:

-   **alternative** for two tailed or one tailed
    -   "two.sided" for a two tailed test,
    -   "greater" for one tailed where the mean of the first variable is greater than the mean of the second variable
    -   "less" for one tailed where the mean of the first variable is greater than the mean of the second variable
    -   `unpaired_t_test_result <- t.test(NonDiabetic, Diabetic, alternative="greater")`
-   **paired** for paired or unpaired
    -   TRUE for paired
    -   FALSE for unpaired
    -   `unpaired_t_test_result <- t.test(NonDiabetic, Diabetic, paired = TRUE)`
-   **mu** mean, for comparison for a one-tailed test